---
title: "Observed Data Prep for Model Analysis - Soils"
author: "PAZ"
date: "31/01/2018"
output: pdf_document
---

```{r, echo=FALSE, message=FALSE, include=FALSE}
Sys.setlocale("LC_ALL", "English")
SAVE = F # Mode
```

# Purpose

Generate BEACH soil calibration data with:

- Time.csv (Julian day marker)
- Composite soils (MonitoringScopeSoils_R.csv)
- Detailed soils (DetailConc.csv & DetailIsotopes.csv)


```{r}
MAC = T
WIN = F
if (MAC) {
  if (WIN){
    path = file.path("C:/Users/DayTimeChunks/Documents/PhD/HydrologicalMonitoring")
    
  } else {
    # path = file.path("/Users/DayTightChunks/Documents/PhD/HydrologicalMonitoring")
    path = file.path("/Users/DayTightChunks/Documents/PhD/hydrological-monitoring")
    time = read.csv2("/Users/DayTightChunks/Documents/PhD/Models/phd-model-master/Analysis/Data/Time.csv")
    time$DayMoYr = as.POSIXct(strptime(time$Date, "%d/%m/%Y", tz="EST"))
  }
} else {
  path = file.path("D:/Documents/these_pablo/Alteckendorf2016/HydrologicalMonitoring")
  time = read.csv2("D:/Documents/these_pablo/Models/BEACH2016/Analysis/Data/Time.csv")
  time$DayMoYr = as.POSIXct(strptime(time$Date, "%d/%m/%Y", tz="EST"))
}
# Lab parameters and field constants
source(file.path(path, "global.R"))
```


# Packages

```{r, message=FALSE}

# Plotting functions
library("scales")
library("tidyr")

library("reshape")
library("zoo") # na.approx()

library("plyr")
library("dplyr")

```

## Working directory

```{r, message=FALSE}
# setwd("D:/Documents/these_pablo/Alteckendorf2016/R")
getwd()
```

# Get composites and convert transect name

```{r}
s = read.csv2(file.path(path, "Data/MonitoringScopeSoils_R.csv"))
s$Transect = as.character(s$Transect)
s$Transect = ifelse(s$Transect == 'T', 'V', s$Transect)
s$DayMoYr = as.POSIXct(strptime(s$Date.Soil, "%d/%m/%Y", tz="EST"))
s = subset(s, s$Wnum > 0)

```

## Merge Transects with Time (Julian Days), ouput as TSS for Nash computation

```{r}

north = subset(s, Transect == "N")
valley = subset(s, Transect == "V")
south = subset(s, Transect == "S")

mean(north$Conc.mug.g.dry.soil, na.rm = T)
mean(valley$Conc.mug.g.dry.soil, na.rm = T)
mean(south$Conc.mug.g.dry.soil, na.rm = T)

mean(north$comp.d13C, na.rm = T)
mean(valley$comp.d13C, na.rm = T)
mean(south$comp.d13C, na.rm = T)

njd = merge(time, north, by = "DayMoYr", all = T)
vjd = merge(time, valley, by = "DayMoYr", all = T)
sjd = merge(time, south, by = "DayMoYr", all = T)

norConc_tss = njd[, c("Jdays", "Conc.mug.g.dry.soil")]
valConc_tss = vjd[, c("Jdays", "Conc.mug.g.dry.soil")]
souConc_tss = sjd[, c("Jdays", "Conc.mug.g.dry.soil")]

norConc_tss$Conc.mug.g.dry.soil = ifelse(is.na(norConc_tss$Conc.mug.g.dry.soil), 
                                         -1e9, norConc_tss$Conc.mug.g.dry.soil)
valConc_tss$Conc.mug.g.dry.soil = ifelse(is.na(valConc_tss$Conc.mug.g.dry.soil), 
                                         -1e9, valConc_tss$Conc.mug.g.dry.soil)
souConc_tss$Conc.mug.g.dry.soil = ifelse(is.na(souConc_tss$Conc.mug.g.dry.soil), 
                                         -1e9, souConc_tss$Conc.mug.g.dry.soil)

norDelta_tss = njd[, c("Jdays", "comp.d13C")]
valDelta_tss = vjd[, c("Jdays", "comp.d13C")]
souDelta_tss = sjd[, c("Jdays", "comp.d13C")]

norDelta_tss$comp.d13C = ifelse(is.na(norDelta_tss$comp.d13C), 
                                1e9, norDelta_tss$comp.d13C)
valDelta_tss$comp.d13C = ifelse(is.na(valDelta_tss$comp.d13C), 
                                1e9, valDelta_tss$comp.d13C)
souDelta_tss$comp.d13C = ifelse(is.na(souDelta_tss$comp.d13C), 
                                1e9, souDelta_tss$comp.d13C)

if (SAVE) {
  write.table(njd, "BEACH_R/north.tss", sep="\t", row.names = F)
  write.table(vjd, "BEACH_R/valley.tss", sep="\t", row.names = F)
  write.table(sjd, "BEACH_R/south.tss", sep="\t", row.names = F)  
  write.table(norConc_tss, "BEACH_R/northConc.tss", sep="\t", row.names = F)
  write.table(valConc_tss, "BEACH_R/valleyConc.tss", sep="\t", row.names = F)
  write.table(souConc_tss, "BEACH_R/southConc.tss", sep="\t", row.names = F)  
  write.table(norDelta_tss, "BEACH_R/northDelta.tss", sep="\t", row.names = F)
  write.table(valDelta_tss, "BEACH_R/valleyDelta.tss", sep="\t", row.names = F)
  write.table(souDelta_tss, "BEACH_R/southDelta.tss", sep="\t", row.names = F)  
}


```

## Get all-composites dataframe for 1:1 calibration graphs

```{r}
# Merge complete dataframe with time
st = merge(time, s, by = "DayMoYr", all = T)
st$IDcal = paste(st$Transect, st$Jdays, sep = "-")

conc_tss = st[, c("Jdays", "Conc.mug.g.dry.soil")]
delta_tss = st[, c("Jdays", "comp.d13C")]

conc_cal = st[, c("Jdays", "Transect", "IDcal", "Conc.mug.g.dry.soil", "Conc.ComSoil.SD")] 
names(conc_cal) <- c("Jdays", "Transect", "IDcal", "ug.g", "ug.g.SD")
conc_cal = subset(conc_cal, !is.na(conc_cal$ug.g))
delta_cal = st[, c("Jdays", "Transect", "IDcal", "comp.d13C", "comp.d13C.SD")]
names(delta_cal) <- c("Jdays", "Transect", "IDcal", "d13C", "d13C.SD")
delta_cal = subset(delta_cal, !is.na(delta_cal$d13C))

if (SAVE) {
  write.table(conc_cal, "BEACH_R/conc_comp_cal.tss", sep="\t", row.names = F)
  write.table(delta_cal, "BEACH_R/d13c_comp_cal.tss", sep="\t", row.names = F)
}
```

# Convert gravimetric to volumetric moisture (used in calibration)

```{r}
grav = read.csv('BEACH_R/theta_grav.csv')
pb_r = read.csv('BEACH_R/pb_range.csv')
pb_r$Date = as.POSIXct(strptime(pb_r$Date, "%d/%m/%Y", tz="EST"))
grav$Date = as.POSIXct(strptime(grav$Date, "%d/%m/%Y", tz="EST"))

pb_merged = merge(pb_r, grav, by = "Date", all = T)
pb_merged = pb_merged[complete.cases(pb_merged$theta_grav),]

pb_merged$theta = pb_merged$theta_grav*pb_merged$pb_ave
pb_merged$theta_min = pb_merged$theta_grav*pb_merged$pb_min
pb_merged$theta_max = pb_merged$theta_grav*pb_merged$pb_max
pb_merged$IDcal = paste(pb_merged$Transect, pb_merged$Jdays, sep = "-")

comp_theta_cal = pb_merged[, c("Jdays", "Transect", "IDcal", "theta")] 


bulk_theta_cal <- comp_theta_cal %>%
  group_by(Jdays) %>%
  dplyr::summarize(theta = mean(theta))

bulk_theta_cal$IDcal = paste("B", bulk_theta_cal$Jdays, sep = "-")

if (SAVE) {
  write.table(comp_theta_cal, "BEACH_R/theta_comp_cal.tss", sep="\t", row.names = F)
  write.table(bulk_theta_cal, "BEACH_R/theta_bulk_cal.tss", sep="\t", row.names = F)
}
```


# Detail concentration and isotopes

## Steps
- group isotope values by ID
- evaluate standard deviations
- remove unlikely delta values based on concentration
- compute means
- merge with concentrations

```{r}
co = read.csv("DetailConc.csv")
is = read.csv("DetailIsotopes.csv")

if (length(is) == 1) {
  is = read.csv("DetailIsotopes.csv", sep = ";")
}

if (length(co) == 1) {
  co = read.csv("DetailConc.csv")
}



# Remove likley GC-MS errors for individual plots > 19 ug/g
co = subset(co, co$ug.g < 19)
length(co$ug.g)
# Remove NA's
is = subset(is, !is.na(d13C))


```

### Group

```{r}
# Same procedure as in Book 06_MassSoils_Composite
siso <- is[, c("ID", "d13C")]
sumIS = ddply(siso, c("ID"), 
              summarise,
              N_detsoil = length(d13C),
              d13C.SD = sd(d13C),
              d13C = mean(d13C)
              )
sumIS = subset(sumIS, N_detsoil > 1 & d13C.SD < 0.8)


length(sumIS$d13C)
```

### Merge conc and isotopes, delete isotope if low concentration (low reliability)

```{r}
ci = merge(co, sumIS, by = "ID", all = T)
ci$DayMoYr = as.POSIXct(strptime(ci$Sampled, "%d/%m/%Y", tz="EST"))
ci = subset(ci, !is.na(ci$DayMoYr))

ci$d13C = ifelse(ci$ug.g < 0.2, NA, ci$d13C)
ci$d13C.SD = ifelse(is.na(ci$d13C), NA, ci$d13C.SD)
```

### Change Talweg's name, merge with time

```{r}
ci$Transect = as.character(ci$Transect)
ci$Transect = ifelse(ci$Transect == 'T', 'V', ci$Transect)

ci$TrPlot = paste(ci$Transect, ci$Plot, sep = "")
ci = ci[, c("DayMoYr", "ID", "Transect", "Plot", "TrPlot", "ug.g", "d13C", "d13C.SD")]
detailed = merge(time, ci, by = "DayMoYr", all = T)
```

# Create plot subsets for plotting against model TSS

```{r}
N1 = subset(detailed, is.na(TrPlot) | TrPlot == "N1")
N2 = subset(detailed, is.na(TrPlot) | TrPlot == "N2")
N3 = subset(detailed, is.na(TrPlot) | TrPlot == "N3")
N4 = subset(detailed, is.na(TrPlot) | TrPlot == "N4")
N5 = subset(detailed, is.na(TrPlot) | TrPlot == "N5")
N7 = subset(detailed, is.na(TrPlot) | TrPlot == "N7")
N8 = subset(detailed, is.na(TrPlot) | TrPlot == "N8")

T4 = subset(detailed, is.na(TrPlot) | TrPlot == "V4")
T5 = subset(detailed, is.na(TrPlot) | TrPlot == "V5")
T7 = subset(detailed, is.na(TrPlot) | TrPlot == "V7")
T8 = subset(detailed, is.na(TrPlot) | TrPlot == "V8")
T9 = subset(detailed, is.na(TrPlot) | TrPlot == "V9")
T10 = subset(detailed, is.na(TrPlot) | TrPlot == "V10")

S11 = subset(detailed, is.na(TrPlot) | TrPlot == "S11")
S12 = subset(detailed, is.na(TrPlot) | TrPlot == "S12")
S13 = subset(detailed, is.na(TrPlot) | TrPlot == "S13")

nor = list(N1, N2, N3, N4, N5, N7, N8)
val = list(T4, T5, T7, T8, T9, T10)
sou = list(S11, S12, S13)

```

## Save TSS detailed

```{r}
if (SAVE) {
  
  counter = 0
  for (d in nor) {
    counter = counter + 1
    transect = "N"
    if (counter < 6) {
      name = paste(transect, as.character(counter), ".tss", sep = "")
    }
    else if (counter == 6) {
      counter = counter + 1
      name = paste(transect, as.character(counter), ".tss", sep = "")
    } else {
      name = paste(transect, as.character(counter), ".tss", sep = "")
    } 
    write.table(d, paste("BEACH_R/", name, sep = ""), sep="\t", row.names = F)
  }
  
  counter = 3
  for (d in val) {
    counter = counter + 1
    transect = "V"
    if (counter < 6) {
      name = paste(transect, as.character(counter), ".tss", sep = "")
    }
    else if (counter == 6) {
      counter = counter + 1
      name = paste(transect, as.character(counter), ".tss", sep = "")
    } else {
      name = paste(transect, as.character(counter), ".tss", sep = "")
    } 
    write.table(d, paste("BEACH_R/", name, sep = ""), sep="\t", row.names = F)
  }
  
  counter = 10
  for (d in sou) {
    counter = counter + 1
    transect = "S"
    name = paste(transect, as.character(counter), ".tss", sep = "")
    write.table(d, paste("BEACH_R/", name, sep = ""), sep="\t", row.names = F)
  }
}

```

# Get all-detailed dataframe for 1:1 calibration graphs 

```{r}
detailed$IDcal = paste(detailed$TrPlot, detailed$Jdays, sep = "-")

conc_det_cal = detailed[, c("Jdays", "Transect", "IDcal", "ug.g")] 
conc_det_cal = subset(conc_det_cal, !is.na(conc_det_cal$ug.g))

delta_det_cal = detailed[, c("Jdays","Transect", "IDcal",  "d13C", "d13C.SD")] 
delta_det_cal = subset(delta_det_cal, !is.na(delta_det_cal$d13C))

if (SAVE) {
  write.table(conc_det_cal, "BEACH_R/conc_det_cal.tss", sep="\t", row.names = F)
  write.table(delta_det_cal, "BEACH_R/d13c_det_cal.tss", sep="\t", row.names = F)
}

```

