---
title: "Observed Data Prep for Model Analysis - Soils"
author: "PAZ"
date: "31/01/2018"
output: pdf_document
---

```{r, echo=FALSE, message=FALSE, include=FALSE}
Sys.setlocale("LC_ALL", "English")

MAC = F
WIN = F

SAVE = F
```

# Purpose

Generate BEACH soil calibration data with:

- Time.csv (Julian day marker)
- Composite soils (MonitoringScopeSoils_R.csv)
- Detailed soils (DetailConc.csv & DetailIsotopes.csv)


```{r}
if (MAC) {
  if (WIN){
    path = file.path("C:/Users/DayTimeChunks/Documents/PhD/HydrologicalMonitoring")
    
  } else {
    # path = file.path("/Users/DayTightChunks/Documents/PhD/HydrologicalMonitoring")
    path = file.path("/Users/DayTightChunks/Documents/PhD/HydroMonitor/.nosync/HydrologicalMonitoring")
    time = read.csv2("/Users/DayTightChunks/Documents/PhD/Models/.nosync/pesti-beach16/Analysis/Data/Time.csv")
    time$DayMoYr = as.POSIXct(strptime(time$Date, "%d/%m/%Y", tz="EST"))
  }
} else {
  path = file.path("D:/Documents/these_pablo/Alteckendorf2016/HydrologicalMonitoring")
  time = read.csv2("D:/Documents/these_pablo/Models/BEACH2016/Analysis/Data/Time.csv")
  time$DayMoYr = as.POSIXct(strptime(time$Date, "%d/%m/%Y", tz="EST"))
}
# Lab parameters and field constants
source(file.path(path, "global.R"))
```


# Packages

```{r, message=FALSE}

# Plotting functions
library("scales")
library("tidyr")

library("reshape")
library("zoo") # na.approx()

library("plyr")
library("dplyr")


```

## Working directory

```{r, message=FALSE}

# setwd("D:/Documents/these_pablo/Alteckendorf2016/R")

# MAC
# setwd("/Users/DayTightChunks/Documents/PhD/Routput/Alteck/R")

# Mac-WIN
# setwd("C:/Users/DayTightChunks/Documents/Models/pesti-beach16/Analysis/Data")
getwd()

```


# Merge Transects with Time (Julian Days)

```{r}
s = read.csv2(file.path(path, "Data/MonitoringScopeSoils_R.csv"))

#q$Date = as.POSIXct(strptime(q$DateCheck, "%d/%m/%Y %H:%M", tz="EST"))
s$DayMoYr = as.POSIXct(strptime(s$Date.Soil, "%d/%m/%Y", tz="EST"))



north = subset(s, Transect == "N")
valley = subset(s, Transect == "T")
south = subset(s, Transect == "S")

mean(north$Conc.mug.g.dry.soil)
mean(valley$Conc.mug.g.dry.soil)
mean(south$Conc.mug.g.dry.soil)

njd = merge(time, north, by = "DayMoYr", all = T)
vjd = merge(time, valley, by = "DayMoYr", all = T)
sjd = merge(time, south, by = "DayMoYr", all = T)

norConc_tss = njd[, c("Jdays", "Conc.mug.g.dry.soil")]
valConc_tss = vjd[, c("Jdays", "Conc.mug.g.dry.soil")]
souConc_tss = sjd[, c("Jdays", "Conc.mug.g.dry.soil")]

norConc_tss$Conc.mug.g.dry.soil = ifelse(is.na(norConc_tss$Conc.mug.g.dry.soil), -1, norConc_tss$Conc.mug.g.dry.soil)
valConc_tss$Conc.mug.g.dry.soil = ifelse(is.na(valConc_tss$Conc.mug.g.dry.soil), -1, valConc_tss$Conc.mug.g.dry.soil)
souConc_tss$Conc.mug.g.dry.soil = ifelse(is.na(souConc_tss$Conc.mug.g.dry.soil), -1, souConc_tss$Conc.mug.g.dry.soil)

norDelta_tss = njd[, c("Jdays", "comp.d13C")]
valDelta_tss = vjd[, c("Jdays", "comp.d13C")]
souDelta_tss = sjd[, c("Jdays", "comp.d13C")]

norDelta_tss$comp.d13C = ifelse(is.na(norDelta_tss$comp.d13C), 0, norDelta_tss$comp.d13C)
valDelta_tss$comp.d13C = ifelse(is.na(valDelta_tss$comp.d13C), 0, valDelta_tss$comp.d13C)
souDelta_tss$comp.d13C = ifelse(is.na(souDelta_tss$comp.d13C), 0, souDelta_tss$comp.d13C)


if (SAVE) {
  write.table(njd, "north.tss", sep="\t", row.names = F)
  write.table(vjd, "valley.tss", sep="\t", row.names = F)
  write.table(sjd, "south.tss", sep="\t", row.names = F)  
  write.table(norConc_tss, "northConc.tss", sep="\t", row.names = F)
  write.table(valConc_tss, "valleyConc.tss", sep="\t", row.names = F)
  write.table(souConc_tss, "southConc.tss", sep="\t", row.names = F)  
  write.table(norDelta_tss, "northDelta.tss", sep="\t", row.names = F)
  write.table(valDelta_tss, "valleyDelta.tss", sep="\t", row.names = F)
  write.table(souDelta_tss, "southDelta.tss", sep="\t", row.names = F)  
}


```

# Merge detail concentration and isotopes

## Steps
- group isotope values by ID
- evaluate standard deviations
- compute means
- merge with concentrations

```{r}

co = read.csv("DetailConc.csv")
is = read.csv("DetailIsotopes.csv")

if (length(is) == 1) {
  is = read.csv("DetailIsotopes.csv", sep = ";")
}


if (length(co) == 1) {
  co = read.csv("DetailConc.csv")
}

is = subset(is, !is.na(d13C))
```

### Group

```{r}
# Same procedure as in Book 06_MassSoils_Composite
siso <- is[, c("ID", "d13C")]
sumIS = ddply(siso, c("ID"), 
              summarise,
              N_detsoil = length(d13C),
              det.d13C = mean(d13C),
              det.d13C.SD = sd(d13C)
              )
sumIS = subset(sumIS, N_detsoil > 1 & det.d13C.SD < 0.8)
```


# Merge conc and isotopes

```{r}
ci = merge(co, sumIS, by = "ID", all = T)
ci$DayMoYr = as.POSIXct(strptime(ci$Sampled, "%d/%m/%Y", tz="EST"))
ci = subset(ci, !is.na(ci$DayMoYr))

```

```{r}
ci$Plot = paste(ci$Transect, ci$Plot, sep = "-")
ci = ci[, c("DayMoYr", "ID", "Plot", "ug.g", "det.d13C", "det.d13C.SD")]
detailed = merge(time, ci, by = "DayMoYr", all = T)
```

# Create plot subsets

```{r}
N1 = subset(detailed, is.na(Plot) | Plot == "N-1")
N2 = subset(detailed, is.na(Plot) | Plot == "N-2")
N3 = subset(detailed, is.na(Plot) | Plot == "N-3")
N4 = subset(detailed, is.na(Plot) | Plot == "N-4")
N5 = subset(detailed, is.na(Plot) | Plot == "N-5")
N7 = subset(detailed, is.na(Plot) | Plot == "N-7")
N8 = subset(detailed, is.na(Plot) | Plot == "N-8")

T4 = subset(detailed, is.na(Plot) | Plot == "T-4")
T5 = subset(detailed, is.na(Plot) | Plot == "T-5")
T7 = subset(detailed, is.na(Plot) | Plot == "T-7")
T8 = subset(detailed, is.na(Plot) | Plot == "T-8")
T9 = subset(detailed, is.na(Plot) | Plot == "T-9")
T10 = subset(detailed, is.na(Plot) | Plot == "T-10")

S11 = subset(detailed, is.na(Plot) | Plot == "S-11")
S12 = subset(detailed, is.na(Plot) | Plot == "S-12")
S13 = subset(detailed, is.na(Plot) | Plot == "S-13")

nor = list(N1, N2, N3, N4, N5, N7, N8)
val = list(T4, T5, T7, T8, T9, T10)
sou = list(S11, S12, S13)

```

## Save detailed

```{r}

if (SAVE) {
  counter = 0
  for (d in nor) {
    counter = counter + 1
    transect = "N"
    if (counter < 6) {
      name = paste(transect, as.character(counter), ".tss", sep = "")
    }
    else if (counter == 6) {
      counter = counter + 1
      name = paste(transect, as.character(counter), ".tss", sep = "")
    } else {
      name = paste(transect, as.character(counter), ".tss", sep = "")
    } 
    write.table(d, paste("BEACH_R/", name, sep = ""), sep="\t", row.names = F)
  }
  
  counter = 3
  for (d in val) {
    counter = counter + 1
    transect = "T"
    if (counter < 6) {
      name = paste(transect, as.character(counter), ".tss", sep = "")
    }
    else if (counter == 6) {
      counter = counter + 1
      name = paste(transect, as.character(counter), ".tss", sep = "")
    } else {
      name = paste(transect, as.character(counter), ".tss", sep = "")
    } 
    write.table(d, paste("BEACH_R/", name, sep = ""), sep="\t", row.names = F)
  }
  
  counter = 10
  for (d in sou) {
    counter = counter + 1
    transect = "S"
    name = paste(transect, as.character(counter), ".tss", sep = "")
    write.table(d, paste("BEACH_R/", name, sep = ""), sep="\t", row.names = F)
  }
}


```

