---
title: "BEACH Formalisms"
header-includes: \usepackage{natbib} \usepackage{mathtools}
output:
  pdf_document:
    citation_package: natbib
    keep_tex: yes
  word_document: default
bibliography: ../../../library.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Note: For word, replace string:
# ${[...comment..]:$.$ <- $.$}  # ignore dots.
# [comment]:$$

```

```{r include=FALSE}
# automatically create a bib database for R packages
# knitr::write_bib(c(.packages(), 'knitr', 'rmarkdown'), 'packages.bib' )
```

# Methodology

## Hydrological Framework

To determine the change in soil moisture content ($\frac{d\theta_i}{dt}$) in the unsaturated zone at each cell *i*, the following hydrological processes are included:

[comment]:$$
\begin{equation}
D\frac{d\theta_i}{dt} = P_i + R_i + \Delta LF_i - Ea_i - Ta_i - DP_i 
\label{eq:hydro} 
\end{equation}
[comment]:$$

where *D* is depth of soil moisture simulation (mm), $\theta$ is soil moisture content (m^3^ m^-3^), *dt* is model time step (day), *P* is precipitation (mm), *R* is runoff (mm), $\Delta LF$ is the difference between lateral inflow and outflow from a given cell (mm), *Ea* is the actual evaporation (mm), *Ta* actual transpiration (mm) and *DP* is the deep percolation (mm).


### Infiltration and Runoff

To calculate infiltration *I*  and surface runoff *R*, soil moisture conditions are determined by following the SCS curve number (BEACH-CN) method defined by the U.S. Soil Conservation Service (SCS, 1972). The method depends on permeability, land use, slope and antecedent moisture conditions. Curve numbers are classified according to three moisture conditions: dry (wilting point - $CN_1$), average moisture ($CN_2$) and wet (field capacity - $CN_3$). 

Typical curve number values for average moisture conditions (i.e. $CN_2$) for various land covers, hydrologic conditions  and soil types at a 5% slope are given in Nietsch et al. [-@Neitsch2009]. $CN_2$ values are adjusted dynamically (i.e. per time step)  by crop cover and leaf area index (LAI) and used to derive the value of $CN_3$ before slope adjustment:

[comment]:$$
\begin{equation}  
CN_3 = CN_2 \cdot \exp [0.00673 \cdot (100-CN_2)] 
\label{eq:CN3}  
\end{equation}
[comment]:$$

The CN2 values are then re-adjusted to account for slope differences such that [@Neitsch2009]:

\begin{equation}  
CN_{2s} = \frac{CN_3 - CN_2}{3} \cdot [1-2 \cdot \exp (-13.86 \cdot slope)] + CN_2 
\label{eq:CN2s}  
\end{equation}

where $CN_{2s}$ is the curve number for average moisture conditions adjusted to the local slope. $CN_1$ values accounting for slope are then calculated as:

\begin{equation}  
CN_{1} = CN_{2s} - \frac{20 \cdot (100 - CN_{2s}}{\Big(100 - CN_{2s} + \exp [2.533-0.0636 \cdot (100 - CN_{2s})]\Big)} 
\label{eq:CN1}  
\end{equation}

Finally, adjustment to $CN_3$ values are calculated by using the original equation and replacing $CN_{2s}$ by $CN_2$.

The run-off equation is given by \cite{Neitsch2009}:

\[
    R = 
\begin{dcases}
     ~0,                                     & P \leq I_a \\
    \frac{(P-I_a)^2}{P-I_a+S},              & P > I_a 
\end{dcases}
\stepcounter{equation}\tag{\theequation}\label{eq:R}
\]


where $I_a$ (mm) is the initial abstraction capacity  of the surface layer, which includes surface storage, interception and infiltration prior to runoff, and typically ranges from 0.05S to 0.2S \citep{Lim2006}. The model adopts the latter of these values as it has provided reliable results for previous rainfall-runoff events. S is the retention parameter after run-off (mm) given as a function of the soil profile water content:

\begin{equation}
S = S_{max} \cdot \Big(1-\frac{SW}{(SW+ \exp[w_1-w_2 \cdot SW]  )} \Big) 
\label{eq:S}  
\end{equation}

where w1 (mm) and w2 (-) are shape coefficients, SW is the soil profile water content excluding the amount of water held in the soil profile at wilting point (mm) such that:

\begin{equation}
SW = max \Big[ \Big\{(\frac{D_{zl} \theta _{zl} + D_1 \theta _1}{D_{zl}+D_1} - \theta _{wp} ) \cdot (D_{zl}+D_1) \Big\},\Big\{0 \Big\} \Big]
\label{eq:SW}
\end{equation}

and $S_{max}$ is the maximum value that the retention parameter (mm) can take such that:

\begin{equation}
S_{max} = 254 \cdot \big(\frac{100}{CN_1} -1 \big)
\label{eq:Smax}  
\end{equation}

Calculation of $w_1$ and $w_2$ further assumes that:

\begin{equation}
S_3= 254 \cdot \big(\frac{100}{CN_3}-1 \big)
\label{eq:S3}
\end{equation}

and that when completely saturated $CN$ = 99 (S = 2.54 mm) such that:

\begin{equation}
w_1 = ln \Big[\frac{FC}{(1-\frac{S_3}{S_{max}}} )-FC \Big]+w_2 \cdot FC
\label{eq:w1}
\end{equation}

\begin{equation}
w_2 = \frac{ln \Big[\frac{FC}{(1-\frac{S_3}{S_{max}}} )-FC \Big]- ln \Big[ \frac{SAT}{(1-\frac{2.54}{S_{max}}}-SAT \Big]}{SAT-FC}
\label{eq:w2}  \end{equation}

where FC is the soil profile water content at field capacity (mm), S3 is the retention parameter (mm) corresponding to field capacity (i.e. CN3) and SAT is the soil profile water content at saturation (mm).  

Infiltration (mm) is then given by:

\begin{equation}
I=P-R	
\label{eq:I}  
\end{equation}

After computation of the runoff equation and infiltration (and in order to compute evapotranspiration, drainage and lateral flow), the cell moisture content for the top layer needs to be updated. This is done by adding infiltration (I) across the soil profile and subtracting the fraction of water content that exceeds the saturation capacity  $\theta_{satex,t}$ such that:

\begin{equation}
\theta_{t+1}=(\frac{\theta_t+I}{D_1} )-\theta_{satex,t}
\label{eq:thetat1}  
\end{equation}

\begin{equation}
\theta_{satex,t}=(\frac{\theta_t+I}{D_1})-\theta_{sat},  ~~~~if~~(\frac{\theta_t+I}{D_1})>\theta_{sat}
\label{eq:thetat}  
\end{equation}

The water content exceeding saturation capacity plus calculated runoff is then added together and reported as actual runoff. The discharge at the outlet is calculated here as a function of time and spatial propagation (i.e. LDD, Runoff). 

### Deep Percolation

Deep percolation (DP) is assumed to be negligible at moisture levels below field capacity. Above this moisture level percolation is given by \cite{Raes2002}:

\begin{equation}
DP_z = D_z \tau_z (\theta_{sat, z} - \theta_{fc, z}) \frac{e^{\theta_z-\theta_{fc,z}}-1}{e^{\theta_{sat, z}-\theta_{fc,z}}-1},~~if~ \theta_z > \theta_{fc, z} 
\label{eq:DP}  
\end{equation}

where $D_z$ is the soil profile depth of layer $z$ and $\tau$ is a dimensionless drainage characteristic given by:


\begin{equation}
\tau = 0.0866 \cdot e^{0.8063 \cdot log_{10}(K_{sat})}, ~~0< \tau \leq 1
\label{eq:tau}  
\end{equation}

where $K_{sat}$ is the saturated hydraulic conductivity.

### Lateral/subsurface flow

Later flow occurs when the soil moisture content exceeds the field capacity and is represented as the difference between inflow and outflow at a cell $j$ for each soil layer $z$ according to \cite{Manfreda2005}:

\begin{equation}
\Delta LF_{j(t)} = \Big( \frac{W_{j} \sum^{N(t)}_{i=1}max[c_{z}(SW_{i}-SW_{fc,i}),~0] }{ \sum^{N(t)}_{i=1} W_{i} } \Big) - max[c_z(SW_j-SW_{fc,j}),~0]
\label{eq:LF}  
\end{equation}

were $c_z$ is the subsurface flow coefficient ($\approx$ 0.25 $d^{-1}$, if $dt$ $\leq$ $1~d$), $SW$ and $SW_{fc}$ are the soil water content at time $t$ and at field capacity for the soil profile [mm], respectively. $N(t)$ is the number of cells $i$ upstream exceeding field capacity and at steepest slope to cell $j$. The topographical wetness index (TWI) $W$, introduced by \cite{Beven1979}, describes the tendency for water to accumulate spatially in regions characterized by a relatively low local slope and a large upstream drainage area. $W_j$ is the TWI at cell $j$ and $W_i$ the cumulative TWI from upstream cells $i$. The TWI is given by:    

\begin{equation}
TWI =ln\Big(\frac{a}{tan \beta} \Big)
\label{eq:Wz}  
\end{equation}

were $a$ is the drainage area per unit contour length (defined by the Local Drain Direction (LDD) network and cell area) and tan $\beta$, the local slope in radians derived from the Digital Elevation Model (DEM).

### Evapotranspiration

To account for evapotranspiration processes the FAO56 reference evaporation rate, $ET_0$ (mm), has been considered and adjusted dynamically according to crop and climate-specific factors. The approach assumes a dual crop coefficient approach appropriate for daily time-step calculations [@Allen1998] and made up of a basal crop coefficient ($K_{cb}$) and a soil water evaporation coefficient ($K_e$). 

Potential evapotranspiration $(ET_p)$ is then given by

\begin{equation}
ET_p=K_c \cdot ET_0
\label{eq:ETp}  
\end{equation}

\begin{equation}
K_c = K_{cb} + K_e
\label{eq:Kc}  
\end{equation}

where $K_{cb}$ varies according to crop-specific development stage. In cases where the mean value for daily relative humidity during the mid- or late-season growth stage ($RH_{min}$\%) differs from 45\% or where wind speed varies by more than 2 m/s the $K_{cb}$ values for mid- and late-season must be adjusted according to:

\begin{equation}
K_{cb}=K_{cb_{mid/end}} + \Big[ 0.04(U_2-2)-0.004(RH_{min}-45) \Big] \Big( \frac{h_{crop}}{3} \Big) ^{0.3}
\label{eq:Kcb}  
\end{equation}

\begin{equation}
K_e= K_{cmax}-K_{cb}
\label{eq:Ke}  
\end{equation}    

where $K_{cb_{mid/end}}$ represent the reference values for sub-humid climate and moderate wind speeds [see @Allen1998]. $U2$ is the wind speed at a height of 2 meters (m/s), $RH_{min}$ is the minimum relative humidity (\%) and $h_crop$ is crop height.
    
$K_e$ is the soil evaporation coefficient and $K_{cmax}$ represents an upper limit to evapotranspiration from cropped surfaces, typically ranges between 1.05 to 1.30 (-) and given by [@Sheikh2009]:  

\begin{equation}
K_{cmax}=max \Big[ \Big\{ K_{cb}+0.05 \Big\} , \Big\{ 1.2 + [0.04(U_2-2)-0.004(RH_{min}-45)] \cdot (\frac{h}{3})^{0.3} \Big\} \Big]
\label{eq:Kcmax}  
\end{equation}

When available soil moisture is limited potential evapotranspiration is reduced (i.e., $ET_a < ET_p$). 

#### Transpiration

To account for potential transpiration processes, water uptake by roots is considered and regulated by atmospheric demand and soil water content. When there is sufficient water in the soil, potential transpiration ($T_p$) equals atmospheric demand [@Sheikh2009]\footnote{Error in Sheikh2009, where $K_{c}f$ should be $K_{cb}$}: 

\begin{equation}
T_p=K_{cb} \cdot ET_0
\label{eq:Tp}  
\end{equation}

Potential transpiration is further subject to root water uptake capacity where the maximum daily uptake $T_{p(z)}$ (mm) at each layer $z$ is given by [@Prasad1988]:

\begin{equation}
T_{p(z)} = 2 \Big( 1- \frac{RD_{z/2} }{RD} \Big) \Big( \frac{RD_z}{RD} \Big) T_p
\label{eq:Tpz}  
\end{equation}

where $RD$ (m) is the total or individual soil layer's rooting depth and $RD_{z/2}$ is the soil depth at the middle of the root extension for layer $z$.    

When soil water is insufficient to meet atmospheric demand, actual transpiration is lower than potential transpiration and given by [@Sheikh2009]:

\begin{equation}
T_{a(z)} = K_s \cdot T_p
\label{eq:Taz}  
\end{equation}

\begin{equation}
K_s= max \Big[ 0, min(1, \frac{\theta_t - \theta_{wp} }{ \theta_c - \theta_{wp} }) \Big]
\label{eq:Ks}  
\end{equation}

\begin{equation}
\theta_c = \theta_{wp} + (1 - p)(\theta_{fc}- \theta_{wp}) 
\label{eq:theta_c}  
\end{equation}

\begin{equation}
p = p_{tab} + 0.04(5-ET_p)
\label{eq:p}  
\end{equation}

where $K_s$ is a transpiration reduction parameter (0-1), which depends on soil water content, $\theta_t$ $(m^3/m^3)$ and the critical soil moisture content $\theta_c$ $(m^3/m^3)$ that defines the transition between unstressed and stressed transpiration rate. $p$ (-) is the fraction of total depletable soil water and $p_{tab}$ the depletion factor (-) for $ET_p \approx 5$ $mm/d$ [@Allen1998, Table no. 22]. 

#### Evaporation

Evaporation is considered only on bare surfaces and assumed to be negligible under plant cover and regulated by atmospheric deman along the first $\approx$ 0.15 m of soil [@Sheikh2009]. Considering the difference between actual ($E_a$, mm/d) and potential evaporation ($E_p$, mm/d) [@Allen1998]:

\begin{equation}
E_p=K_e \cdot ET_0
\label{eq:Ep}  
\end{equation}

\begin{equation}
E_a=K_r \cdot E_p
\label{eq:Ea}  
\end{equation}

where $K_r$ is an evaporation reduction coefficient (-) given by:

\begin{equation}
K_r = \frac{ \theta_t - \theta_{dr} }{ \theta_{fc} - \theta_{dr} }
\label{eq:Kr}  
\end{equation}

where $\theta_t$ is soil moisture ($m^3/m^3$) and $\theta_{dr}$ is the moisture ($m^3/m^3$) of air-dry soil\footnote{Note: In model, $\theta_{dr} = 0.33 \cdot \theta_{wp}$ (Sheikh et al., 2009)}. 


### Crop Growth

#### Height

Height is determined as a function of the stage of crop development. 

[comment]:$$
\begin{equation} 
    h = 
\begin{dcases}
     ~~~~~0~,                                     & t \leq t_{sow-date} \\
     \frac{~h_{max} \cdot~(t-t_{sow-date})}{\Delta t_{}} ,     & t \leq t_{sow-date} \\
    \frac{lnf}{-5\cdot10^{-5}},              & f > 0 
\end{dcases}
\end{equation}
[comment]:$$



- Root depth


### Soil Temperature

Temperture is an important parameter regulating chemical and biological processes. To simulate its evolution across time and space this model follows [@Neitsch2009], where the average daily soil temperature ($^{\circ}$C) at the center of each soil layer $z$ is given by:

\begin{equation}
T_{soil_z}(t) = l \cdot T_{soil_, k,_z}(t-1) + (1.0 - l) \cdot \big[df \cdot [\overline{T}_{AAir} - T_{ssurf}]+T_{ssurf}]
\label{eq:tempSoil} 
\end{equation}

where $l$ is a lag coefficient (0 - 1.0) regulating the influence of the previous day's $(t-1)$ temperature, $\overline{T}_{AAir}$ is the average annual air temperature and $T_{ssurf}$ is the soil surface temperature at time $t$. The depth factor $df$ is given by,

\begin{equation}
df = \frac{zd}{zd+exp(-0.867-2.078 \cdot zd)}
\label{eq:df} 
\end{equation}

where $zd$ is the ratio of the depth at the center of the soil layer $D_{z/2}$ (mm) to the damping depth $dd$ (mm) such that:
\begin{equation}
zd = \frac{D_{z/2}}{dd}
\label{eq:zd} 
\end{equation}

The damping depth $dd$ is a function of the maximum damping depth $dd_{max}$ (mm) and a soil water scaling factor $\varphi$ (-).

\begin{equation}
dd = dd_{max} \cdot exp \Big[ln\Big(\frac{500}{dd_{max}} \Big) \cdot \Big(\frac{1-\varphi}{1+\varphi}\Big)^2\Big]
\label{eq:dd} 
\end{equation}

\begin{equation} 
dd_{max} = 1000 + \frac{2500\rho_b}{\rho_b+686 \cdot exp(-5.63\rho_b)}
\label{eq:ddmax} 
\end{equation}

\begin{equation} 
\varphi = \frac{SW}{(0.356-0.144\rho_b) \cdot D_{k}}
\label{eq:varphi} 
\end{equation}

where $\rho_b$ is the soil bulk density $(mg~m^{-3})$ and SW is water content $(mm~H_2O)$ in the soil profile $D_z~(mm)$. 

The soil surface temperature $T_{ssurf}$ in eq. \ref{eq:tempSoil}, is a fucntion of the previous day's soil temperature, amount of ground cover captured by a crop factor $bcv$ and the temperature of bare soil $T_{bare}$, such that:

\begin{equation} 
T_{ssurf} = bcv \cdot  T_{soil, k, t-1} + (1-bcv) \cdot T_{bare}
\label{eq:Tssurf} 
\end{equation}

\begin{equation} 
bcv = \frac{CV}{(CV+exp(7.563-1.297 \cdot 10^{-4} \cdot CV))}
\label{eq:Tssurf} 
\end{equation}

[comment]:$$
\begin{equation} 
    CV = 
\begin{dcases}
     ~~~~~0~,                                     & f \leq 0 \\
    \frac{lnf}{-5\cdot10^{-5}},              & f > 0 
\end{dcases}
\end{equation}
[comment]:$$


\begin{equation} 
f=1-exp(-\mu*LAI)
\label{eq:f} 
\end{equation}

where $\mu$ is light-use efficiency of the crop $(kg~ha^{-1}~m^2~MJ^{-1})$.

## Pesticide Fate 

### Pesticide distribution

The concentration of pesticide within the soil system ($g/L$) along each layer $z$ is given as a function of its partitioning across the soil's gaseous, aqueous and adsorbed phases by:

\begin{equation}
C_{tot_z} = \theta_{gas_z}(t)C_{gas_z}(t) + \theta_z(t)C_{aq_z}(t) + \rho_{b_z}(t)C_{ads_z}
\label{eq:conc_tot}  
\end{equation}

where $\theta_{gas_z}(t) = \theta_{sat_z}(t) - \theta_{z}(t)$, $C_{gas_z}$ is the concentration in the gas phase $(g/L)$, $C_{aq_z}$ the concentration in aqueous phase ($g/L$) and $C_{ads_z}$ the concentration in adsorbed phase ($g/Kg~soil$) with the bulk soil density $\rho_{b_z}$ in $Kg/L$.

Converting to mass $(g)$ based on the cell area ($A_i$, $m^2$) and model layer depth ($D_z$, $mm$):

\begin{equation}
M_{tot_z} = A_iD_z\theta_{gas_z}(t)C_{gas_z}(t) + A_iD_z\theta_z(t)C_{aq_z}(t) + A_iD_z\rho_{b_z}(t)C_{ads_z}
\label{eq:mass_tot}  
\end{equation}

simplifying to the respective fractions,

\begin{equation}
M_{tot_z} = V_{gas_z}(t)C_{gas_z}(t) + V_{H_2O}(t)C_{aq_z}(t) + M_{soil}(t)C_{ads_z}
\label{eq:mass_tot_simple}  
\end{equation}

and substituting phase concentrations for their equivalent in $C_{aq}$ according to eq. \ref{eq:kd} and eq. \ref{eq:henry} yields,

\begin{equation}
M_{tot_z} = V_{gas_z}C_{aq_z}(t)/K_H + V_{H_2O}(t)C_{aq_z}(t) + M_{soil}(t)K_dC_{aq_z}(t)
\label{eq:mass_tot_sub}  
\end{equation}

Solving for $C_{aq}$ $(g/L~H_2O)$ with $K_d$ units in $(L/Kg)$,

\begin{equation}
C_{aq} = \frac{M_{tot_z} }{  V_{gas_z}(t)/K_H + V_{H_2O}(t) + M_{soil}(t)K_d }
\label{eq:mass_tot_conc_aq1}  
\end{equation}

\begin{equation}
C_{aq} = \frac{M_{tot_z} }{ A_iD_z \Big( \theta_{gas_z}(t)/K_H + \theta_{aq}(t) + \rho_{b_z}(t)K_d \Big)}
\label{eq:mass_tot_conc_aq2}  
\end{equation}

and substituting the retardation factor from eq. \ref{eq:retard_linear},

\begin{equation}
C_{aq} = \frac{M_{tot_z} }{ A_iD_z \Big( \theta_{gas_z}(t)/K_H + \theta_{aq}(t)R_z(t) \Big)}
\label{eq:mass_tot_conc_aq2}  
\end{equation}



Aqueous concentration is obtained with the following function. Note that the input parameter, `mass`, may be equal to the total concentration or either the heavy or light fractions, whichever output is required by the user.


```{python}

def getConcAq(model, layer, theta_sat, mass,
              sorption_model="linear", gas=True):
    # Note that p_b (g/cm3) x k_d (L/Kg) -> unit-less
    if layer == 0:
        depth = model.z0
        theta_layer = model.theta_z0
    elif layer == 1:
        depth = model.z1
        theta_layer = model.theta_z1
    elif layer == 2:
        depth = model.z2
        theta_layer = model.theta_z2

    if sorption_model == "linear":
        # Retardation factor
        retard_layer = 1 + (model.p_b * model.k_d) / theta_layer
    else:
        print("No sorption assumed, Ret. factor = 1")
        retard_layer = 1  # No retardation.

    if gas:  # Leistra et al., 2001
        theta_gas = max(theta_sat - theta_layer, scalar(0))
        conc_aq = mass / ((cellarea() * depth) *  # m2 * mm = L
                          (theta_gas / model.k_h +
                           theta_layer * retard_layer))  # ug/L cell volume
    else:  # No gas phase
        # Whelan, 1987
        conc_aq = (mass / (cellarea() * depth * theta_layer * retard_layer)) 

    return conc_aq

```

An analogous re-arrangement is applied to obtained the concentration in gas $C_{gas}$ $(g/L~air)$ and solid (adsorbed) $C_{ads}$ $(g/Kg~soil)$ phases:

\begin{equation}
C_{g} = \frac{C_{aq}}{K_H} 
\label{eq:conc_gas}  
\end{equation}

\begin{equation}
C_{ads} = \frac{M_{tot_z} }{ A_i\cdot D_z \Big( \theta_{gas_z}(t)/(K_H \cdot K_d) + \theta_{aq}(t)/K_d + \rho_{b_z}(t)\Big)}
\label{eq:mass_tot_conc_aq2}  
\end{equation}

To obtain the adsorbed concentration $C_{ads}$ $(g/Kg~soil)$, the following function is used:

```{python}
def getConcAds(model, layer, theta_sat, mass, gas=True):
    # mass / Kg soil
    if layer == 0:
        depth = model.z0
        theta_layer = model.theta_z0
    elif layer == 1:
        depth = model.z1
        theta_layer = model.theta_z1
    elif layer == 2:
        depth = model.z2
        theta_layer = model.theta_z2

    if gas:
        theta_gas = max(theta_sat - theta_layer, scalar(0))
        # [mass pest/Kg soil]
        conc_ads = mass / ((cellarea() * depth) *
                           (theta_gas / (model.k_h * model.k_d) +
                            theta_layer / model.k_d +
                            model.p_b))
    else:
        print("No implementation without gas available")
        raise NotImplementedError

    return conc_ads
```



### Sorption

#### Linear sorption 

The partition of pesticide concentrations into the dissolved $C_{aq}$ ($g/L$) and adsorbed phases $C_{ads}$ $(g/Kg)$ is determined by the pesticide dissociation coefficient $K_d$ $(L/Kg)$:

\begin{equation}
K_d = \frac{ C_{ads} }{ C_{aq}  }
\label{eq:kd}
\end{equation} 

\begin{equation}
K_d = K_{oc} \cdot f_{oc}
\label{eq:kd}
\end{equation} 

where $K_{oc}$ $(L/Kg)$ is the pesticide octanol partition coefficient and $f_{oc}$ is the fraction of organic carbon (-) in soil. The retardation factor $R_z$ (-) is given by:

\begin{equation}
R_z(t) = 1 + \frac{ \rho_{b_z}(t) \cdot K_d }{ \theta_z(t) }
\label{eq:retard_linear}
\end{equation} 


### Volatilization

```{r include=FALSE}
# Note: Consider only where evaporation occurs (i.e. 0.15 m depth), but only first layer (i.e. < 15cm)
```

Pesticide volatilization follows Leistra \textit{et al.} [-@Leistra2001], where a boundary air layer is conceptualized through which pesticide difusses before escaping into the atmosphere. The thickness ($d_a$, $m$) of this layer, assumed to be equivalent to the topmost soil layer's thickness or mixing layer, regulates the transport resistance ($r_a$, $d/m$) such that:

\begin{equation}
r_a(t) = \frac{ d_a }{ D_a(t) }
\label{eq:resistance_air}
\end{equation} 

where $D_a$ ($m^2/d$) is the diffusion coefficient in air for Metolachlor at the observed environmental temperature and adjusted relative to the reference diffusion coefficient ($D_{a,r}$, $m^2/d$) as:


\begin{equation}
D_a(t) = \Big(\frac{ T(t) }{ T_r } \Big)^{1.75} D_{a,r}
\label{eq:Da}
\end{equation} 

where $T$ and $T_r$ are the environmental temperature at time $t$ and at the reference temperature at 293.15$^\circ$K, respectively.

The total volatilization is given by the flux across the air layer boundary ($J_{v,air}$) and the flux across the topmost soil layer ($J_{v,soil}$) such that: 

\begin{equation}
J_{v,air}(t) = - \frac{ C_{gas,top}(t) - C_{air}(t) }{ r_a }
\label{eq:Jva}
\end{equation} 

\begin{equation}
J_{v,soil}(t) = - \frac{ C_{gas,z_0}(t) - C_{gas,top}(t) }{ r_s }
\label{eq:Jvs}
\end{equation} 

where $C_{gas,top}$ ($mg/m^3$) is the concentration in gas phase at the soil surface, $C_{air}$ ($mg/m^3$) the concentration in air, $C_{gas,z_0}$ ($mg/m^3$) the concentration in gas phase at the center of the uppermost soil layer and $r_s$ ($d/m$) the diffusion resistance across the topmost soil layer and given by:

\begin{equation}
r_s(t)= \frac{ 0.5 D_z }{ D_{rdiff,g}(t) }
\label{eq:resistance_soil}
\end{equation} 

To calculate the relative diffusion ($D_{rdiff,gas}$, $m^2/d$) the model provides two options. Under option 1 [@Millington1960],

\begin{equation}
D_{rdiff,gas} = \frac{ D_a(t) \Big(\theta_{gas_z}(t)\Big)^a}{ \Big(\theta_z(t) \Big)^b }
\label{eq:d_rdiff1}
\end{equation} 

where Jin and Jury [-@Jin1996] recommend that $a=2$ and $b=2/3$. Under option 2 [@Currie1960],

\begin{equation}
D_{rdiff,gas} = D_a(t) \Big(a\Big) \Big( \theta_{gas_z}(t)  \Big)^b
\label{eq:d_rdiff2}
\end{equation} 

where Bakker \textit{et al.} [-@Bakker1987] recommend $a=2.5$ and $b=3$ for moderately aggregated plough layers of loamy soils and humic sandy soils [@Leistra2001].

Finally, it is assumed that flux across both layer boundaries is equivalent ($J_{v,soil} = J_{v,air}$) [@Leistra2001]. Considering pesticide concentration in air to be negligble ($C_{air} = 0$), the concentration at the soil surface is:

\begin{equation}
C_{gas,top}(t) = \frac{r_a}{(r_a + r_s)} C_{gas,z_{0}(t)}
\label{eq:conc_gas_top}
\end{equation} 

The gas concentration in the soil layer is related to the dimensionless Henry constant ($K_H$), where:

\begin{equation}
C_{gas,z_0}(t) = C_{aq,z_0}(t) K_H  
\label{eq:henry}
\end{equation} 

Substituting eq. \ref{eq:conc_gas_top} into eq. \ref{eq:Jva} yields the mass flux lost to the atmosphere ($g/m^2d$):

\begin{equation}
J_{v,air} = - \frac{C_{gas,z_0}}{(r_a + r_s)}
\label{eq:Jva_final}
\end{equation} 


To obtain the pesticide mass volatilized at each time step, the following function is defined:

```{python}
def getVolatileMass(model, temp_air, theta_sat, mass, frac,
                    rel_diff_model="option-1", sorption_model="linear",
                    gas=True, isotopes=True, ):
    # Volatilize only during peak volatilization time i.e., first 24 hrs, @Prueger2005.
    theta_layer = model.theta_z0
    theta_gas = max(theta_sat - theta_layer, scalar(0))
    # Convert to m (needed for final mass computation on cell basis)
    depth_m = model.z0 * 1 / 10 ** 3
    # Air boundary layer, assumed as 2m high
    thickness_a = scalar(1.0)  # m
    # Diffusion coefficient in air (cm^2/s); https://www.gsi-net.com
    #  D_ar (metolachlor) = 0.03609052694,  at reference Temp., in Kelvin, D_a,r)
    diff_ar = 0.03609052694 * 86400.0 * 1.0 / 10 ** 4  # m2/d
    # Diffusion coefficient adjusted to air Temp. in Kelvin, D_a
    diff_a = ((temp_air + 273.15) / 293.15) ** 1.75 * diff_ar  # m2/d

    if rel_diff_model == "option-1":
        # Millington and Quirk, 1960 (in Leistra, 2001, p.48)
        # a,b parameters: Jin and Jury, 1996 (in Leistra, 2001)
        diff_relative_gas = (diff_a * theta_gas ** 2 /
                             theta_sat ** (2 / 3))  # m2/d
    elif rel_diff_model == "option-2":
        # Currie 1960 (in Leistra, 2001)
        # a,b parameters: Baker, 1987 (in Leistra, 2001)
        diff_relative_gas = diff_a * 2.5 * theta_gas ** 3  # m2/d
    else:
        print("No appropriate relative diffusion parameter chosen")
        diff_relative_gas = diff_a  # m2/d
    # Transport resistance through air (r_a) and soil (r_s) layer
    r_a = thickness_a / diff_a  # d/m
    r_s = (0.5 * depth_m) / diff_relative_gas  # d/m

    conc_aq = getConcAq(model, 0, theta_sat, mass,
                        sorption_model=sorption_model, gas=gas, isotopes=isotopes)
    # Convert ug/L to ug/m3, as will be multiplying by cell's area in m2
    conc_aq *= 10 ** 3  # ug/L * 10^3 L/m3
    conc_gas = conc_aq / model.k_h  # ug/L air
    volat_flux = (conc_gas / (r_a + r_s)) * cellarea()  # ug/day
    return volat_flux
```


### Run-off Loss

#### \textit{Non-uniform mixing-layer-model-runoff (nu-mlm-ro)}

Multiple models are available to simulate mass transfer to overland flow. The first of these models, the \textit{Non-uniform mixing-layer-model-runoff} (nu-mlm-ro) is adapted from Ahuja and Lehman, 1983 [see @Shi2011, eq. 1 and p. 1217] and given by:

\begin{equation}
\frac{\partial (EDI \theta C_m)}{\partial t} = -RO \beta C_m
\label{eq:nu-mlm-ro}
\end{equation} 

\begin{equation}
\beta = e^{(-bz)}
\label{eq:beta-nu-mlm}
\end{equation} 

where the Effective Depth of Interaction (EDI) refers to the mixing layer depth $(mm)$, $\theta$ is soil moisture $(L/L)$, RO is run-off $(mm)$ and $C_m$ is concentration in the mixing layer $(g/L)$. The paramater $b$ is a calibration constant (assuming, $1 \ge b > 0$) and where $z$ is the depth of the simulated top-soil layer. In this model, $\beta$ accounts for an exponential decrease in the ability of overland flow to mix with soil water as depth increases. 

#### \textit{Non-uniform mixing-layer (nu-mlm)}

The adaptation above of the original model by Ahuja and Lehman replaces precipitation by $RO$, and thus considers the ability of rainfall water to mix with soil water instead. To test the original formulation, this second model is also made available as \textit{Non-uniform mixing layer model} (nu-mlm). 

#### \textit{Distributed mixing-layer model (d-mlm)}

The \textit{Distributed mixing-layer model} considered is adapted from Havis [-@Havis1992], and mass transfer to overland flow based on a mass flux coefficient $(K_L$, $mm/day)$. The change in mass in the overland flow (and consequently in the mixing layer) is given by:

\begin{equation}
\frac{\partial (h_{runoff} C_{runoff}) }{\partial t} + \frac{\partial(qC_{runoff}) }{\partial x} = K_L (C_m - C_{runoff})
\label{eq:d-mlm-of}
\end{equation} 

where $h_{runoff}$ is overland flow height (mm), $q$ is overland flow discharge rate per unit width (mm$^2$/day).   

Instead of considering $K_L$ as calibration parameter, the model defines $K_L$ (mm/day) as the ratio of the mass flux ($\varrho$, mg/mm$^2$ day) from the soil surface to overland flow and the solute concentration difference between overland flow and the surface soil (i.e., $K_L = \varrho /(C_m - C_{runoff})$). Note that due to small concentations in overland flow, it may be assumed that $C_{runoff} \approx 0$. For laminar flow, based on Bennett and Myers [-@Bennet1982], $K_L$ is given by: 

\begin{equation}
K_L = 0.664 \frac{D_w}{L}Re^{1/2}S_c^{1/3}
\label{eq:K_L}
\end{equation} 

\begin{equation}
Re = \frac{\rho v L}{\mu}
\label{eq:Re}
\end{equation} 

\begin{equation}
S_c = \frac{ \mu}{\rho D_w}
\label{eq:S_c}
\end{equation} 

where the $D_w$ is the solute diffusivity (cm$^2$/s), $Re$ is the Reynolds number (-) and $S_c$ is the Schmidt number (-). Parameters include the cell length (L, mm), the dynamic viscocity of water ($\mu$ at 25 $^\circ C$, 8.9e-03 g/cm sec), the soil bulk density ($\rho$, g/cm$^3$) and the runoff velocity ($v$, mm/day) or amount of runoff generated in each cell per day. 

To obtain the mass lost to run-off the following function is defined:

```{python}
def getRunOffMass(model, theta_sat, precip, runoff_mm,
                  mass, frac,
                  transfer_model="simple-mt", sorption_model="linear",
                  gas=True, isotopes=True):
    # Aqueous concentration
    conc_aq = getConcAq(model, 0, theta_sat, mass,
                        sorption_model=sorption_model, gas=gas, isotopes=isotopes)

    if transfer_model == "simple-mt":
        mass_ro = conc_aq * runoff_mm * cellarea()
    elif transfer_model == "nu-mlm-ro":
        # non-uniform-mixing-layer-model-runoff (nu-mlm-ro)
        # Considers a decrease in effective transfer as mixing layer depth increases
        # Adapted from Ahuja and Lehman, 1983 in @Shi2011,
        # Adaptation replaces Precip by Runoff amount.
        b = 1  # [mm] Calibration constant, 1 >= b > 0 (b-ranges appear reasonable).
        # As b decreases, mass transfer increases, model.z0 in mm
        mass_ro = (runoff_mm * cellarea()) * exp(-b * model.z0) * conc_aq 
    elif transfer_model == "nu-mlm":
        # non-uniform-mixing-layer-model (nu-mlm)
        # Original from Ahuja and Lehman, 1983 in @Shi2011
        b = 1  # [mm] Calibration constant, 1 >= b > 0 (b-ranges appear reasonable).
        # As b decreases, mass transfer increases, model.z0 in mm
        mass_ro = (precip * cellarea()) * exp(-b * model.z0) * conc_aq
    elif transfer_model == "d-mlm":
        # distributed mixing-layer-model (d-mlm)
        # Adapted from Havis et al., 1992, and
        # taking the K_L definition for laminar flow from Bennett and Myers, 1982.
        mass_ro = getKfilm(model, runoff_mm) * cellarea() * conc_aq  
    else:
        print("Run-off transfer model not stated")
        return None

    return mass_ro
```

To obtain $K_L$ the following function is defined and implemented when `transfer_model = 'd-mlm'` is declared in `getRunOffMass()` above:   

```{python}
def getKfilm(model, runoffvelocity):
    """
    Note: Model uses run-off (mm) per day (i.e. timestep) as runoff velocity.
    Chemical parameter source:
    http://www.gsi-net.com/en/publications/gsi-chemical-database/single/377.html
    """
    # Dynamic viscosity of water (\mu) @25 Celsius = 8.9e-04 [Pa s]
    #   1 Pa = 1 N/(m s^2) = 1 Kg/(m s^2)
    #   Convert to g/(cm s): dyn_visc = 8.9e-03 [g/cm s]
    dyn_visc = 8.9e-03  # [g/cm s] @25 degrees, [@Shi2011]:\mu

    # Solute diffusivity in water (D_w)
    # Metolachlor = 5.0967719112e-006 (cm2 / s)
    diff_solute = 5.0967719112e-006  # [cm2 / s], [@Shi2011]:D_w
    Sc = dyn_visc / (model.p_b * diff_solute)  # (-) Schmidt number, [@Shi2011]:S_c

    # Reynolds number (dimensionless), 86400s = 1 day
    cell_length = 2 * 10 ** 3  # mm
    # Reynolds (Re), [-] (Shi et al., 2011)
    re = (model.p_b * 1 / 10 ** 2 * runoffvelocity * cell_length) / (dyn_visc * 86400)
    kl = (0.664 * ((diff_solute * 86400 * 10 ** 2) / cell_length) *
          re ** (float(1) / 2) * Sc ** (float(1) / 3)) 
    return kl  # mm/day
```


### Vertical mass flux (i.e., leaching)

Vertical flux can be computed differently across soil layers. Under the first approach, and only for the uppermost layer, the model follows McGrath [-@Mcgrath2008]:

\begin{equation}
C_{z_0,aq}(t+1) = C_{z_0,aq}(t) exp \Big( \frac{ -P(t) }{ \theta_{z_0}(t) \cdot R_{z_0}(t) \cdot D_{z_0} } \Big) 
\label{eq:conc_mcgrath}
\end{equation} 

The mass leached ($g$) is thus given by:

\begin{equation}
M_{z_0,lch}(t) =  D_{z_0} \cdot A_i  \Big(\theta_{z_0}(t)C_{z_0,aq}(t)- \theta_{z_0}(t+1) C_{z_0,aq}(t+1) \Big) 
\label{eq:leached_mcgrath}
\end{equation} 

where $A$ is the area ($m^2$) for each cell $i$. 

Under the second approach, available on all layers, mass leached is proportional to the aqueous concentration in percolated water such that: 

\begin{equation}
M_{z,lch}(t) = DP_z(t) \cdot C_{z,aq }(t) \cdot A_i
\label{eq:leached_prop}
\end{equation} 

```{python}
def getLeachedMass(model, layer, theta_sat,
                   water_flux,
                   theta_after_percolate,
                   mass,
                   sorption_model=None,
                   leach_model=None, gas=True, isotopes=True):
    if layer == 0:
        depth = model.z0
        theta_layer = model.theta_z0
    elif layer == 1:
        depth = model.z1
        theta_layer = model.theta_z1
    elif layer == 2:
        depth = model.z2
        theta_layer = model.theta_z2

    # Aqueous concentration
    conc_aq = getConcAq(model, layer, theta_sat, mass,
                        sorption_model=sorption_model, gas=gas, isotopes=isotopes)

    if sorption_model == "linear":
        # Retardation factor
        retard_layer = scalar(1) + (model.p_b * model.k_d) / theta_layer
    else:
        print("No sorption assumed, Ret. factor = 1")
        retard_layer = scalar(1)  # No retardation.

    if leach_model == "mcgrath":
        if layer == 0:
            conc_aq_new = conc_aq * exp(-water_flux / (theta_layer * retard_layer * depth))
            mass_aq = conc_aq * (theta_layer * depth * cellarea())
            mass_aq_new = conc_aq_new * (theta_after_percolate * depth * cellarea())
            mass_leached = mass_aq - mass_aq_new
            if mapminimum(mass_aq_new) < 0:
                print("Error in Leached Model")
        else:
            # McGrath not used in lower layers,
            # as formulation accounts for rainfall impact
            mass_leached = conc_aq * water_flux * cellarea()
            mass_aq = conc_aq * (theta_layer * depth * cellarea())
            mass_aq_new = mass_aq - mass_leached
            if mapminimum(mass_aq_new) < 0:
                print("Error in Leached Model")
    else:
        mass_leached = conc_aq * water_flux * cellarea()
        mass_aq = conc_aq * (theta_layer * depth) * cellarea()
        mass_aq_new = mass_aq - mass_leached
        if mapminimum(mass_aq_new) < 0:
            print("Error in Leached Model")

    return mass_leached
```


### Lateral mass flux

Later mass flux is proportional to the aqueous concentration in lateral water flow. It is obtained by adapting eq. \ref{eq:LF}, where the difference between mass loss in upstream cells $i$ and mass loss at cell $j$ is:

\begin{equation}
\Delta LMF_{j,z} = \sum^{N(t)}_{i=1} M_{loss, i(t)} - M_{loss,j(t)}
\label{eq:LMF}  
\end{equation}

The mass gain at cell $j$ is given by the mass loss from upstream cells $i$ contributing to downstream cells and given by:

\begin{equation}
\sum^{N(t)}_{i=1} M_{loss, i(t)} = \frac{W_{j} \sum^{N(t)}_{i=1}max[ C_{i,aq}\cdot\Big(c_{z}(SW_{i}-SW_{fc,i})\Big),~0] }{ \sum^{N(t)}_{i=1} W_{i} }
\label{eq:LFMgain} 
\end{equation}
Loss at cell $j$ is then given by:

\begin{equation}
M_{loss,j(t)} = C_{j,aq}\cdot\Big(c_{z}(SW_{j}-SW_{fc,j})\Big)
\end{equation}
 
Simplfying for the relative wetness index at cell $j$ in eq. \ref{eq:LFMgain}, we obtain:

\begin{equation}
W_{j/i} = \frac{W_{j}}{\sum^{N(t)}_{i=1} W_{i} } 
\end{equation}

```{python}
def getLatMassFlux(model, layer, theta_sat, theta_fcap,
                   mass, sorption_model='linear', gas=True, isotopes=True):
    if layer == 0:
        depth = model.z0
        theta_layer = model.theta_z0
        c = model.c1
    elif layer == 1:
        depth = model.z1
        theta_layer = model.theta_z1
        c = model.c1
    elif layer == 2:
        depth = model.z2
        theta_layer = model.theta_z2
        c = model.c2

    # Aqueous concentration
    conc_aq = getConcAq(model, layer, theta_sat, mass,
                        sorption_model=sorption_model, gas=gas, isotopes=isotopes)

    # W(j/i)
    rel_wetness = model.wetness / accuflux(model.ldd_subs, model.wetness)

    # Cell mass loss/gain (to update only mass)
    mass_loss = max(conc_aq * (c * (depth * theta_layer - depth * theta_fcap)), scalar(0))
    mass_gain = rel_wetness * accuflux(model.ldd_subs, mass_loss)
    net_mass_latflux = mass_gain - mass_loss

    return {
        'mass_loss': mass_loss,
        'mass_gain': mass_gain,
        'net_mass_latflux': net_mass_latflux
    }
```


## Degradation

### Photodegradation

...to code and document.

```{r include=FALSE}
# Note: Only first 24 hrs or first 5 mm cummulative rainfall?
```


### Biodegradation

Biodegradation in the soil is assumed to occur only on the dissolved and sorbed fractions. Assuming a first-order rate law, the change in mass due to degradation is given by:

\begin{equation}
\frac{dM_{tot_z}}{dt} = -k_{b,aq}(V_{H_2O_z(t)}C_{aq_z}(t))-k_{b,ads}(M_{soil_z}C_{ads_z}(t))
\label{eq:dM_bio}
\end{equation}

where $k_{b}$ $(d^{-1})$ is the biodegradation rate constant in the aquoeous $(aq)$ and sorbed soil sites $(ads)$. Considering degradation on each phase respectively equation \ref{eq:dM_bio} can subdivided and written in integrated for concentrations as:

\begin{equation}
C_{aq_z}(t+1) = C_{aq_z}e^{-k_{b,aq} \Delta t} 
\label{eq:Maq_t1}
\end{equation}

\begin{equation}
C_{ads_z}(t+1) = C_{ads_z}e^{-k_{b,ads} \Delta t}
\label{eq:Mads_t1}
\end{equation}

where $C_{aq}$ and $C_{ads}$ are the contamimant concentrations in liquid and sorbed sites, respectively. Updating for total contaminant mass for all fractions:

\begin{equation}
M_{tot_z}(t+1) = V_{gas_z}(t)C_{gas_z}(t) + V_{H_2O}(t)C_{aq_z}(t+1) + M_{soil}(t)C_{ads_z}(t+1)
\label{eq:Mtot_t1}
\end{equation}

The biodegradation rate constant for the aqueous phase $k_{b,aq}$ is given by:

\begin{equation}
k_{b,aq} = \frac{ln(2)}{t_\frac{1}{2}} 
\label{eq:k_b}
\end{equation} 

where $t_\frac{1}{2}$ is the reference half-life (d) of the contaminant considered. The degradation in the sorbed fraction is assumed to be a fraction of that in the aqueous phase such that:

\begin{equation}
k_{b,ads} = s_f \cdot k_{b,aq}   
\label{eq:k_bads}
\end{equation} 

where $0 \leq s_f \leq 1$ $(-)$ is an adjustment parameter for degradation rate. Although a decrease in degradation rates with increasing depth reltive to top soils may be associated to a decrease in microbial activity **(Rodríguez-Cruz et al., 2006; Si et al., 2009)**, currently no depth variation is considered. However, this could be easily implemented by assuming an exponential decrease in the degradation rate constant as a function of depth such that: 

\begin{equation}
k_{{b,aq}_z} = k_{b,aq}\cdot exp(-\gamma \cdot D_{z/2}) 
\label{eq:gamma}
\end{equation} 

where $\gamma$ $(-)$ is a calibration parameter $0 \leq \gamma \leq 1$ and $D_{z/2}~(m)$ the depth to center of layer $z$. 

Adapting the reference half-life to temperature and moisture changes the half-life becomes:

\begin{equation}
t_\frac{1}{2}=t_\frac{1}{2}^{ref}\cdot F_T \cdot F_\theta
\label{eq:DT50}	
\end{equation} 

where $t_\frac{1}{2}^{ref}$ is the half-life (days) at the reference moisture and temperature. $F_T$ and $F_\theta$ are factor changes in degradation rates associated to temperature and moisture conditions, following \cite{Schroll2006} and the Arrhenius equation, respectively. To account for the influence of water content across a range of saturation conditions $F_\theta$ is given by:

[comment]:$$
\begin{equation} 
    F_\theta = 
\begin{dcases}
     ~~~~~0~,                 & \theta_t \leq 0.5 \cdot \theta_{wp} \\
    \Big(\frac{ \theta_t- 0.5 \cdot \theta_{wp} } { \theta_{fc} - \theta_{wp} }\Big)^{\beta_{\theta} },    & \theta_t <  \theta_{fc} \\
     ~~~~~1~,                 & \theta_t > \theta_{fc} 
\end{dcases}
\end{equation}
[comment]:$$

\begin{equation} 
F_T = e^{(E_a/R)(1/T_{ref} - 1/T)}
\end{equation} 

where T is the temperature in soil converted to Kelvin. $T_{ref}$ is the reference temperature at which the experimental half-life is reported (i.e. typically 20 $^{\circ}$C). $E_a$ is the activation energy generally equal to 54,000 (KJ mol$^{-1}$) and R is the gas constant 8.314 (J mol$^{-1}$ K$^{-1}$).

To compute  mass degradation, the following method is used:

```{python}
def getMassDegradation(model, layer,
                       theta_sat, theta_fcap, theta_wp,
                       mass, frac="L", sor_deg_factor=1,
                       sorption_model="linear", gas=True):
    # Get mass applied if top-soil
    if layer == 0:
        theta_aq_layer = model.theta_z0
        temp_layer = model.temp_z0_fin
        depth = model.z0
    elif layer == 1:
        theta_aq_layer = model.theta_z1
        temp_layer = model.temp_z1_fin
        depth = model.z1
    elif layer == 2:
        theta_aq_layer = model.theta_z2
        temp_layer = model.temp_z2_fin
        depth = model.z2

    # F_Theta_1
    theta_factor = ifthenelse(
        theta_aq_layer <= 0.5 * theta_wp, scalar(0),
        ifthenelse(theta_aq_layer <= theta_fcap,
                   (((theta_aq_layer - 0.5 * theta_wp) / (
                       theta_fcap - theta_wp)) ** scalar(model.beta_moisture)),
                   scalar(1)))
    # Temperature factor in biodegradation
    # F_T_1
    temp_factor = exp((model.act_e / model.r_gas) *
                      (1 / (model.temp_ref + 273.15) - 1 / (model.temp_air + 273.15)))

    # Half-life as a function of temperature and moisture
    dt_50 = max(model.dt_50_ref * theta_factor * temp_factor, scalar(0))
    # dt_50 = max(model.dt_50_ref)

    # Convert to degradation constant
    # Deg in dissolved phase
    k_b = ifthenelse(dt_50 > 0, ln(2) / dt_50,
                     scalar(0))  
    # Deg in sorbed phase (now assumed equal)
    k_bs = k_b * sor_deg_factor

    # Step 0 - Obtain species concentration (all phases)
    conc_aq = getConcAq(model, layer, theta_sat, mass,
                        sorption_model=sorption_model, gas=gas)  # mass/L
    conc_ads = getConcAds(model, layer, theta_sat, mass, gas=gas)  # mass/g soil
    conc_gas = conc_aq / model.k_h

    mass_aq = conc_aq * (theta_aq_layer * depth * cellarea())
    mass_ads = conc_ads * (depth * cellarea() * model.p_b)  # pb = g/cm3

    # Step 1 - Degrade phase fractions
    # First order degradation kinetics
    theta_gas = max(theta_sat - theta_aq_layer, scalar(0))
    if frac == "H":
        conc_aq_new = conc_aq * exp(-1 * model.alpha_iso * k_b * scalar(model.jd_dt))
        conc_ads_new = conc_ads * exp(-1 * model.alpha_iso * k_bs * scalar(model.jd_dt))
    else:  # Same for total conc as for "L"
        conc_aq_new = conc_aq * exp(-1 * k_b * scalar(model.jd_dt))
        conc_ads_new = conc_ads * exp(-1 * k_bs * scalar(model.jd_dt))

    # Step 2 - Convert to mass (i.e., after degradation in each phase)
    mass_aq_new = conc_aq_new * (theta_aq_layer * depth * cellarea())
    mass_ads_new = conc_ads_new * (model.p_b * depth * cellarea())  # pb = g/cm3
    mass_gas = conc_gas * (theta_gas * depth * cellarea())
    mass_tot_new = mass_aq_new + mass_ads_new + mass_gas
    mass_deg_aq = mass_aq - mass_aq_new
    mass_deg_ads = mass_ads - mass_ads_new

    return {"mass_tot_new": mass_tot_new,
            "mass_deg_aq": mass_deg_aq,
            "mass_deg_ads": mass_deg_ads}
```


### Isotope treatment

To obtain the masses of heavy $M^h$ and light isotopes $M^l$, the isotope signature $\delta ^{13}C$ is used where:

\begin{equation} 
\delta ^{13}C [\textperthousand] = \Big(\frac{R_{smp} - R_{std}}{R_{std}}\Big)\cdot 1000
\end{equation}


\begin{equation} 
R_{smp} = \frac{M^h}{M^l} 
\end{equation}

\begin{equation} 
R_{std} = 11237.2 \cdot 10^{-6}
\end{equation}


and where the total mass $M_{tot} = M^h + M^l$, we obtain:

\begin{equation} 
R_{std}\Big(\delta ^{13}C/1000+1 \Big)= \frac{M_{tot}-M^l}{M^l}
\end{equation}

\begin{equation} 
M^l = \frac{M_{tot}}{1+R_{std}(\delta ^{13}C/1000+1)}
\end{equation}


\begin{equation} 
M^h = M_{tot} - M^l 
\end{equation}





# References 

