---
title: "BEACH Formalisms"
header-includes: \usepackage{mathtools} \usepackage{natbib}
output:
  pdf_document:
    citation_package: natbib
    keep_tex: yes
  word_document: default
bibliography: library.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Note: For word, replace string:
# ${[...comment..]:$.$ <- $.$}  # ignore dots.

```

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(.packages(), 'knitr', 'rmarkdown'), 'packages.bib' )
```



# Methodology

## Hydrological Framework

To determine the change in soil moisture content ($\frac{d\theta_i}{dt}$) in the unsaturated zone at each cell *i*, the following hydrological processes are included:

[comment]:$$
\begin{equation}
D\frac{d\theta_i}{dt} = P_i + R_i + \Delta LF_i - Ea_i - Ta_i - DP_i 
\label{eq:hydro} 
\end{equation}
[comment]:$$

where *D* is depth of soil moisture simulation (mm), $\theta$ is soil moisture content (m^3^ m^-3^), *dt* is model time step (day), *P* is precipitation (mm), *R* is runoff (mm), $\Delta LF$ is the difference between lateral inflow and outflow from a given cell (mm), *Ea* is the actual evaporation (mm), *Ta* actual transpiration (mm) and *DP* is the deep percolation (mm).