---
title: "Volatilization Alternative explanation"
header-includes: \usepackage{mathtools} \usepackage{natbib}
output:
  pdf_document:
    citation_package: natbib
    keep_tex: yes
  word_document: default
bibliography: library.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Note: For word, replace string:
# ${[...comment..]:$.$ <- $.$}  # ignore dots.
# [comment]:$$
```

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(.packages(), 'knitr', 'rmarkdown'), 'packages.bib' )
```



Transport processes, including volatilization, leaching, lateral mass flux are considered sequentially before degradation and direclty after application. Sorption is considered within each process. 

## Volatilization


Volatilization is assumed only to take only place on the application day, and follows \cite{Leistra2001} by considering mass flux across two boundary layers. Mass flux across an air boundary $J_{v,a}$ $(\mu g/ m^2d)$ and mass flux across the top soil layer $J_{v,s}$ $(\mu g/ m^2d)$ and given by:

$$.
J_{v,air} = \frac{-(C_{g,0(t)}-C_{air})}{r_{air(t)}}
$$

$$.
J_{v,soil} = \frac{-(C_{g,z0}-C_{g,0})}{r_{soil(t)}}
$$
where, $r_a$ $(d/m)$ is the air boundary resistance and $r_s$ $(d/m)$ the top soil layer resistance before the pesticide reaches the soil surface from the center of the cell layer $(0.5 \cdot D_{z0})$. Concentration gradients $(\mu g/m^3)$ are considered based on the concentration at the soil surface $C_{g,0}$, the air boundary layer $C_{air}$ and at the center of top soil layer.  

Resistance across the boundary air and soil layers are given respectively by:

$$.
r_{air(t)} = \frac{D_{z,air}}{D_{a(t)}} 
$$


$$.
r_{soil(t)} = \frac{0.5 \cdot D_{z0}}{D_{diff,g}} 
$$

where $D_{z,air}$ $(m)$ is the boundary air layer assumed to be equal to the the first soil layer $(k=z0)$ thickness and the pesticide diffusion coefficient $D_a$ $(m^2/d)$, which is dependent on the daily air temperature $T$ (in $K$) as:

$$.
D_{a(t)} = \Big( \frac{T}{T_r}\Big)^{1.75}D_{a,r}
$$
and where $D_{a,r}$ $(m^2/d)$, is the diffussion coefficient in air at reference temperature $T_r$. Two options are available to compute the relative diffussion coefficient $D_{diff,g}$ $(m^2/d)$, which is adjusted to the available pore air space $\theta_{z0,air(t)} = \theta_{z0,sat}-\theta_{z0(t)}$. The first option is given by \citep{Millington1960, Jin1996}:

$$.
D_{diff,g(t)} = \frac{D_{a(t)}\cdot (\theta_{z0,air(t)})^2}{(\theta_{z0,sat})^{2/3}}
$$

The second option is given by \citep{Currie1960, Bakker1987}:

$$.
D_{diff,g(t)} = 2.5 \cdot D_{a(t)}\cdot (\theta_{z0,air(t)})^3
$$

During each process, sorption extent is computed (as shown above) to obtain the concentration in aqueous $C_{z0,aq}$  $(\mu g/L)$ and gas $C_{z0,g}$ $(\mu g/L)$ solutions. It is assumed that the diffusion flux from the top soil layer is equal to the diffusion flux across the bounadry air layer and that concentration in air $C_{air}=0$, yielding:

\begin{equation}
C_{g,0} = \frac{r_{air(t)}}{r_{air(t)} + r_{soil(t)}} \cdot C_{g,z0}
\label{eq:Cgas0} 
\end{equation}


Substituting eq. \ref{eq:Cgas0} into \ref{eq:Jvair} ... 

The volatilized flux out of the top layer is given by:

...this has been documented in BEACH Methodoogy!!

